---
title: "Midterm Exam"
author: "Yuelin Jiang"
date: "11/2/2020"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(rstanarm)
library(bayesplot)
```

## Instruction

This is your midterm exam that you are expected to work on it alone.  You may NOT  discuss any of the content of your exam with anyone except your instructor. This includes text, chat, email and other online forums.  We expect you to respect and follow the [GRS Academic and Professional Conduct Code](http://www.bu.edu/cas/files/2017/02/GRS-Academic-Conduct-Code-Final.pdf). 

Although you may NOT ask anyone directly, you are allowed to use external resources such as R codes on the Internet.  If you do use someone's code, please make sure you clearly cite the origin of the code.

When you finish, please compile and submit the PDF file and the link to the GitHub repository that contains the entire analysis.


## Introduction

In this exam, you will act as both the client and the consultant for the data that you collected in the data collection exercise (20pts).  Please note that you are not allowed to change the data. The goal of this exam is to demonstrate your ability to perform the statistical analysis that you learned in this class so far.  It is important to note that significance of the analysis is not the main goal of this exam but the focus is on the appropriateness of your approaches.

### Data Description (10pts)

Please explain what your data is about and what the comparison of interest is.  In the process, please make sure to demonstrate that you can load your data properly into R.

ANSWER:
The data is count data of motorcycles passed at an intersection given the same amount of time (3 minutes). My interest of question is, whether the number of motorcycles with passengers are different between noontime and evening time. The analysis should take into account of the total motorcycles passed during this time.
```{r }
motor <- read.csv("motorcycles.csv")
head(motor)
motor$passTotal <- motor$paseenger2orMore + motor$passenger1

mean(motor$passTotal)
sd(motor$passTotal)
```

### EDA (10pts)

Please create one (maybe two) figure(s) that highlights the contrast of interest.  Make sure you think ahead and match your figure with the analysis.  For example, if your model requires you to take a log, make sure you take log in the figure as well.

```{r }

par(mfrow = c(1,2))
ggplot(data = motor)+
  geom_point(aes(x= log(total), y = log(passTotal) , color = factor(time) ))
```

### Power Analysis (10pts)

Please perform power analysis on the project.  Use 80% power, the sample size you used and infer the level of effect size you will be able to detect.  Discuss whether your sample size was enough for the problem at hand.  Please note that method of power analysis should match the analysis.  Also, please clearly state why you should NOT use the effect size from the fitted model.

```{r }

```


### Modeling (10pts)

Please pick a regression model that best fits your data and fit your model.  Please make sure you describe why you decide to choose the model. Also, if you are using GLM, make sure you explain your choice of link function as well.

ANSWER: 
```{r }

fit <- stan_glm(formula = passTotal ~ factor(time) , family = (link = "log"), offset = log(total), data = motor, refresh=0 )
summary(fit, digits = 4)
```

### Validation (10pts)

Please perform a necessary validation and argue why your choice of the model is appropriate.  

```{r }
post.fit = posterior_predict(fit2)
ppc_dens_overlay( y = motor$passTotal, yrep=post.fit[1:100,])
plot(fitted(fit2),resid(fit2),pch=20) 
curve(sqrt(x),add=T)
ppc_error_scatter_avg(y=motor$passTotal,yrep=post.fit)
# sim = posterior_predict(fit2,draws=1000)
# ppc_dens_overlay(motor$passTotal,sim[1:100,])
```


### Inference (10pts)

Based on the result so far please perform statistical inference to compare the comparison of interest.

```{r }

```


### Discussion (10pts)

Please clearly state your conclusion and the implication of the result.


### Limitations and future opportunity. (10pts)

Please list concerns about your analysis.  Also, please state how you might go about fixing the problem in your future study.


### Comments or questions
If you have any comments or questions, please write them here.

